---
title: "NCAA GSR Parts 2 and 3"
author: "Joy Lindsay"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(utHelpR)
library(here)
library(scales)
library(readxl)
library(janitor)
library(knitr)
```

bringing in the sql from Edify

```{r}
ipeds_graduation_df <- utHelpR::get_data_from_sql_file('ipeds_graduation.sql', 'edify', 'project')
```

bringing in the Excel sheet given to us from Athletics - Note: the file needs to be named just like the file given if you rename the file you need to rename it here as well.

```{r}
gsr_2018_cohort_raw <- read_excel(here("sensitive", "GSR 2018 Cohort 2025.xlsx"))
```

Cleaning up the Excel file to make a cleaner data frame

```{r}
gsr_2018_cohort <- gsr_2018_cohort_raw %>% 
  clean_names()
```

```{r}
# Declare Variables ####
v_cohort_start_term_id <- '201840'
v_cohort_description <- 'First-Time Freshman'
v_full_time_part_time_code <- 'F'
```

# Question 2

```{r warning=FALSE}
ncaa_grad_sum_df <- ipeds_graduation_df %>%
  #filter(cohort_code == "FTFB201640") %>%  # we could just filter by the cohort code
  filter(cohort_start_term_id == v_cohort_start_term_id,
         cohort_desc == v_cohort_description,
         full_time_part_time_code == v_full_time_part_time_code) %>%
  mutate(count_entered = 1) %>% 
  mutate(count_graduated = if_else(!is.na(graduation_date) & ipeds_award_level_code == 5, 1, 0)) %>%
  mutate(count_exclusion = if_else(is_exclusion, 1, 0)) %>%
  select(student_id,count_entered, is_athlete, count_graduated, count_exclusion, 
         ipeds_race_ethnicity, gender_code, is_enrolled) %>% 
  mutate(count_athlete = if_else(is_athlete, 1, 0)) %>% 
  group_by(ipeds_race_ethnicity, gender_code) %>% 
  summarise(across(.cols = starts_with("count"),  ~ sum(.x, na.rm = FALSE)))

```

## Function to make it look like the pdf

```{r}
sum_ncaa_grad <- function(input_df, gender_code){
   gender_code <- rlang::enquo(gender_code) 
  
output_df <- input_df %>% 
  filter(gender_code == {{gender_code}}) %>% 
  mutate(RaceEthnicity = case_when(
           ipeds_race_ethnicity == 'Non-Resident Alien' ~ 1,
           ipeds_race_ethnicity == 'Hispanic' ~ 2,
           ipeds_race_ethnicity == 'American Indian/Alaskan' ~ 3,
           ipeds_race_ethnicity == 'Asian' ~ 4,
           ipeds_race_ethnicity == 'Black/African American' ~ 5,
           ipeds_race_ethnicity == 'Hawaiian/Pacific Islander' ~ 6,
           ipeds_race_ethnicity == 'White' ~ 7,
           ipeds_race_ethnicity == 'Multiracial' ~ 8,
           ipeds_race_ethnicity == 'Unspecified' ~ 9 )) %>% 
  arrange(RaceEthnicity) %>% 
  select(-c(2, 6:7)) %>%
  rename_with(~str_remove(., "count_"), everything()) %>% 
  adorn_totals()

return(output_df)
}
```

Passing the data frame through the function to make two reports, one for Men and the other for Women.

```{r}
men <- sum_ncaa_grad(ncaa_grad_sum_df, "M")

women <- sum_ncaa_grad(ncaa_grad_sum_df, "F")
```

Using Kable to clean it up a bit.

```{r}
kable((men), caption = "Men")

kable((women), caption = "Women")
```

# Question 3

```{r}
ncaa_grad_df <- ipeds_graduation_df %>%
  mutate(count_graduated = if_else(!is.na(graduation_date) & ipeds_award_level_code == 5, 1, 0)) %>%
  mutate(count_exclusion = if_else(is_exclusion, 1, 0)) %>%
  select(student_id, count_graduated, count_exclusion, ipeds_race_ethnicity, 
         is_enrolled, cohort_start_term_id, cohort_desc, cohort_code)
```

## unfiltered

```{r}
gsr_2018_cohort_grad_rate_unfiltered <- gsr_2018_cohort %>% 
  mutate(student_id = str_pad(student_id, side = "left", pad = "0", 8)) %>% 
  left_join(ncaa_grad_df, by = c("student_id" = "student_id"))
```

## filtered by

```{r}
gsr_2018_cohort_grad_rate <- gsr_2018_cohort_grad_rate_unfiltered %>% 
   filter(cohort_start_term_id == v_cohort_start_term_id,
          cohort_desc == v_cohort_description)
```

This will make a new spreadsheet in the sensitive folder.

```{r}
write_csv(gsr_2018_cohort_grad_rate, here::here('sensitive', '2018_cohort_output.csv'))
```
