---
title: "NCAA Graduation Success Rate Report"
author: "Joy Lindsay"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Libraries

```{r echo=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(utHelpR)
library(here)
#library(IPEDSuploadables)
library(scales)
```

### Pull data from Edify using sql and filter on variables defined above.

```{r}
graduation_rate_raw <- utHelpR::get_data_from_sql_file('graduation_rate.sql', 'edify', 'project')

headcount_census_groups_raw <- utHelpR::get_data_from_sql_file('headcount_census_groups.sql', 'edify', 'project')

twelve_month_hc_raw <- utHelpR::get_data_from_sql_file('twelve_month_hc.sql', 'edify', 'project')
```

### Define Variables

#### Variable Definitions:

-   **v_unit_id** - Unit ID of the institution.
-   **v_cohort_start_term_id** - The start term of the cohort.
-   **v_cohort_description** - The cohort description *i.e (First-Time Freshman, Transfer)*.
-   **v_full_time_part_time_code** - Full or part time cohort status.

```{r echo=TRUE}
# Declare Variables ####
# Graduation Rates
v_unit_id <- '230171'
v_cohort_start_term_id <- '201840'
v_cohort_description <- 'First-Time Freshman'
v_full_time_part_time_code <- 'F'
```

### Graduation Rate Munge

This data frame transforms the graduation rate data to match the criteria needed for calulating the IPEDS graduation rate.

```{r echo=TRUE, warning=TRUE, message=TRUE}
gr_munge_df <- graduation_rate_raw %>%
  filter(cohort_start_term_id == v_cohort_start_term_id,
         cohort_desc == v_cohort_description,
         full_time_part_time_code == v_full_time_part_time_code) %>%
  mutate(Unitid = v_unit_id,
         EnteringProgramType = if_else(cohort_degree_level_code == 'BA', 3, 1),
         CurrentProgramType = case_when(ipeds_award_level_code %in% c('1A', '1B') ~ 1,
                                        ipeds_award_level_code %in% c('2', '3') ~ 2,
                                        ipeds_award_level_code == '5' ~ 3
                                        ),
         Completed150 = if_else((( ipeds_award_level_code %in% c('1A', '1B') & days_to_graduate <= 365 ) |
                                    # graduated with a mid credit hour cert within 1.5 years
                                    ( ipeds_award_level_code == '2' & days_to_graduate <= 547 ) |
                                    # graduated with associate within 3 years 
                                    ( ipeds_award_level_code == '3' & days_to_graduate <= 1095 ) |
                                    # graduated with bachelors within 6 years
                                    ( ipeds_award_level_code == '5' & days_to_graduate <= 2190 )),1,0, missing = 0),
         CompletedFourYears = if_else(ipeds_award_level_code == '5' & days_to_graduate <= 1460, 1, 0, missing = 0),
         CompletedFiveYears = if_else(CompletedFourYears != 1 & ipeds_award_level_code == '5' & days_to_graduate <= 1825, 1, 0, missing = 0),
         IsExclusion = if_else(Completed150 != 1  & is_exclusion, 1, 0),
         transfer_out_calc = if_else((Completed150 == 1 | IsExclusion == 1), FALSE, is_transfer_out),
         IsTransferOut = if_else(transfer_out_calc, 1, 0, missing =  0),
         is_still_enrolled_calc = if_else((Completed150 == 1 | IsTransferOut == 1 | IsExclusion == 1), FALSE, is_enrolled),
         IsStillEnrolled = if_else(is_still_enrolled_calc, 1, 0, missing =  0),
         RaceEthnicity = case_when(
           ipeds_race_ethnicity == 'Non-Resident Alien' ~ 1,
           ipeds_race_ethnicity == 'Hispanic' ~ 2,
           ipeds_race_ethnicity == 'American Indian/Alaskan' ~ 3,
           ipeds_race_ethnicity == 'Asian' ~ 4,
           ipeds_race_ethnicity == 'Black/African American' ~ 5,
           ipeds_race_ethnicity == 'Hawaiian/Pacific Islander' ~ 6,
           ipeds_race_ethnicity == 'White' ~ 7,
           ipeds_race_ethnicity == 'Multiracial' ~ 8,
           ipeds_race_ethnicity == 'Unspecified' ~ 9
         ),
         Sex = case_when(gender_code == 'M' ~ 1,
                         gender_code == 'F' ~ 2),
         GenderDetail = case_when(gender_code == 'M' ~ 1,
                                  gender_code == 'F' ~ 2),
         PellGrant = if_else(is_pell_awarded, 1, 0),
         DirectLoan = if_else(is_pell_awarded != 1 & is_subsidized_loan_awarded, 1, 0),
         IsAthlete = if_else(is_athlete, 1, 0),
         IsVeteran = if_else(is_veteran, 1, 0)
         )

```


Bachelor's or equivalent degree-seeking subcohort:

First Time Full Time Bachelor Seeking Freshman:

```{r}
bach_seeking_cohort <- gr_munge_df %>% 
  filter(cohort_degree_level_code == 'BA') %>% 
  select(cohort_start_term_id, cohort_code_desc, ipeds_award_level_code, is_athlete, gender_code, is_graduated, ipeds_race_ethnicity)
```

Completers of bachelor's or equivalent degrees:

```{r}
bach_complete_rate <- bach_seeking_cohort %>%
  mutate(is_graduated = if_else(is.na(is_graduated), FALSE, is_graduated)) %>% 
  group_by(ipeds_award_level_code, is_athlete) %>% 
  summarise(headcount = n()) 



  
```

Demographics - Academic Year 2024-25 (Summer 24, Fall 24, Spring 25 Census)

Full Time Undergraduate Baccalaureate Degree Seeking Students

```{r}

demographics_2425 <- twelve_month_hc_raw %>%
  filter(term_id %in% c('202430', '202440', '202520'), 
         is_degree_seeking == TRUE,
         level_id == 'UG',
         full_time_part_time_code == 'F',
         str_starts(primary_degree_id, "B")) %>% 
  select(student_id, is_degree_seeking, gender_code, full_time_part_time_code, level_id, primary_degree_id) %>%
  distinct(student_id, .keep_all = TRUE) %>% 
  group_by(gender_code) %>% 
  summarise(count = n())








```

